### 2024-01-09 manuscript.Rmd
initiate manuscript doc

---
title: A GIS based workflow for Ecosystem Condition Accounting - an example using spatially biased indicators
author:
  - name: Anders Lorentzen Kolstad
    email: anders.kolstad@nina.no
    affiliation: Norwegian Institute for Nature Research
    correspondingauthor: true
    footnote: 1
  - name: Bob Security
    email: bob@example.com
    affiliation: Another University
  - name: Cat Memes
    email: cat@example.com
    affiliation: Another University
    footnote: 2
  - name: Derek Zoolander
    email: derek@example.com
    affiliation: Some Institute of Technology
    footnote: 2
address:
  - code: Some Institute of Technology
    organization: Big Wig University
    addressline: 1 main street
    city: Gotham
    state: State
    postcode: 123456
    country: United States
  - code: Another University
    organization: Department
    addressline: A street 29
    postcode: 2054 NX
    city: Manchester,
    country: The Netherlands
footnote:
  - code: 1
    text: "This is the first author footnote."
  - code: 2
    text: "Another author footnote."
abstract: |
  This is the abstract.

  It consists of two paragraphs.
keywords: 
  - SEEA EA
  - alien species
  - ...
journal: "Ecological Indicators"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: mybibfile.bib
linenumbers: false
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
---

# Introduction
SEEA EA principles

The need for local field data

Low data availability

Fast turn around with automated workflows

4 types of indicator

# Material and Methods

In this study we focused on open mire ecosystems in Norway.
For this ecosystem we have an ecosystem delineation map produced using a  
a deep learning model which was trained on field verified ground truth data and and satellite imagery. 
This model had a XX% precision when tested against independent field data.

The reference state defining ecosystems in optimal or pristine condition is defined as one with little or no human influence, with a climate as in the period 1961-1990 and a native species similar as today (no mammoths allowed).

We then identified two _concepts_ that define open mire ecosystems in optimal condition:

- Native species dominate
- No physical damage caused by human activities

To describe these two concepts we use five condition variables from the same data set (ref).
The five variables are:

1. Alien plant cover
1. 7SE
1. 7TK
1. PR..
1. PR..

The data set from a standardised field survey of nature types.
Each locality is scored on a range of variables relevant for describing the state and quality of nature.
The surveys are comissioned with the goal of producing rel spatially

Variable 1 was planned to be developed into a single indicator (i.e. a normalised variable) with the same name.
Variables 2-5 describe very related aspects as so they where planned to be combined into a composite indicator called antropogenic disturbance to soil and vegetation so ADSV or short.

Alien plant cover was assigned to ECT class B1a - Compositional state characteristics
ADSV was assigned to ECT class A1 - physical ...

National indicator and for about 4 municipalities as a case study

Validation/sensitivity analyses (varying HIA resolution and minimum _n_?)

Schematic GIS workflow

Aggregation

# Results
Data availability/coverage

Validation

National/regional indicator - do we have enough data?

Municipal case study

# Discussion

# Conclusion

# Credit authorship contribution statement {-}

# Declaration of Competing Interest {-}

# Acknowledgements {-}

# Data availability {-}

# References {-}