### 2024-01-14 R/scrap/bayesian inference.qmd
ways to estimate uncertainty

---
title: "Bayesian inference"
format: html
editor: visual
---

## Bayesian inference

I want to test how I can get best estimates of the population means with a distribution around it to describe the uncertainty. If I just use the sample variance, then with very low sample sizes, like *n =* 1 or 2, I can get very low variance just by chance, making the estimates seem very robust, when in fact it is not.

If I use bayesian inference over a flat prior and using the s.dev. from the national data set (closer to the true sd.dev. than the sub-population sd.dev.), then perhaps I can get what I want?

```{r, warning=F, message=F}
library(Bolstad)
library(tidyverse)
```

Our variable is discrete, 1 to 7.

```{r}
var <- seq(1,7, by=1) 
```

We want to estimate the mean of the population.

The probability density function for the mean (mu) will be normally distributed and truncated at 1 and 7.

Let's generate some example population samples.

```{r}
random1 <- 2
random2 <- c(sample(var, 2, replace = T))
random10 <- c(sample(var, 10, replace = T))

full <- c(sample(var, 1000, replace = T))
```

`Bolstad::normdp` assumes the samples comes from a normal distribution with a known st.dev.

I will take the st.dev from the entire national data set.

```{r}
national.sd <- sd(full)
```

We want to use a uniform prior.

Here's the posterior distribution for mu when we have just *n* = 1

```{r}
pd1 <- normdp(x = random1,
       sigma.x = national.sd,
       n.mu = 50) # default 
```

The plot is truncated at 4. We can see this better if we extract the values manually and make our own plot

```{r}
x1 <- pd1[["param.x"]]
y1 <- pd1[["posterior"]]
df <- data.frame(x1, y1)
df |> 
  ggplot(aes(x = x1, y = y1))+
  geom_area(fill="orange", alpha=.5)+
  xlim(c(0,7))
```

The possible value of mu is restricted. We have to define it.

Mu is actually not discrete. Just the data are.

```{r}
mu <- seq(1, 7, length.out=100)
```

```{r}
pd2 <- normdp(x = random1,
       sigma.x = national.sd,
       n.mu = 50, # not having an effect
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))

```

OK, but there is practically zero change of mu being 7. We only have one sample. Shouldn't there be a wider spread of probabilities?

Actually, the probability that mu is 7 is calculated to be

```{r}
min(pd2$posterior)
```

So not zero, but still very small. And if the st.dev is 2, then perhaps this is correct.

```{r}
dnorm(7, mean=random1, sd=national.sd)
```

This is also a small number, although not identical.

```{r}
pd10 <- normdp(x = random10,
       sigma.x = national.sd,
       n.mu = 50, # not having an effect
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))
```

Lets try some more realistic samples.

```{r}
sampleHigh2 <- round(runif(2, 5,7))
sampleHigh5 <- round(runif(5, 5,7))
sampleHigh10 <- round(runif(10, 5,7))
sampleHigh20 <- round(runif(20, 5,7))
```

```{r, output=F}
pd_h2 <- normdp(x = sampleHigh2,
       sigma.x = national.sd,
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))

pd_h5 <- normdp(x = sampleHigh5,
       sigma.x = national.sd,
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))

pd_h10 <- normdp(x = sampleHigh10,
       sigma.x = national.sd,
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))

pd_h20 <- normdp(x = sampleHigh20,
       sigma.x = national.sd,
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))
```

```{r}
data.frame(x = mu,
                 "n2" = pd_h2$posterior,
                 "n5" = pd_h5$posterior,
                 "n10" = pd_h10$posterior,
                 "n20" = pd_h20$posterior) |>
  pivot_longer(!x, 
               names_to = "n") |>
  ggplot(aes(x = x, y = value, fill = n))+
  geom_area(alpha=.5)+
  theme_bw()

```

This figure is wrong. The data is correct to begin with, but plotting goes wrong.