### 2024-01-15 R/scrap/bayesian inference.qmd
wgt_mean

---
title: "Bayesian inference"
format: html
editor: visual
---

## Mean of population

I want to test how I can get best estimates of the population means with a distribution around it to describe the uncertainty. If I just use the sample variance, then with very low sample sizes, like *n =* 1 or 2, I can get very low variance just by chance, making the estimates seem very robust, when in fact it is not.

Our variable is discrete, 1 to 7.

```{r}
var <- seq(1,7, by=1) 
```

We have a national data set with thousands of samples (*n* \> 1000). From this we want to get the mean for local regions (the data is georeferenced).

If I use bayesian inference over a uniform prior and using the st.dev. from the national data set (closer to the true st.dev. than the sub-population st.dev.), then perhaps I can get what I want?

I will use the Bolstad package for the inference.

```{r, warning=F, message=F}
library(Bolstad)
library(tidyverse)
library(ggridges)
```

We want to estimate the mean of the population.

The probability density function for the mean (mu) will be normally distributed and truncated at 1 and 7.

Let's generate some example population samples.

```{r}
random1 <- 2
random2 <- c(sample(var, 2, replace = T))
random10 <- c(sample(var, 10, replace = T))

full <- c(sample(var, 1000, replace = T))
```

`Bolstad::normgcp` assumes the samples comes from a normal distribution with a known st.dev.

I will take the st.dev from the entire national data set.

```{r}
national.sd <- sd(full)
```

We want to use a uniform prior.

Here's the posterior distribution for mu when we have just *n* = 1

```{r}
pd1 <- normgcp(x = random1,
       sigma.x = national.sd,
       density = "uniform",
       n.mu = 50) # default 
```

The plot goes beyond the possible range of \[1-7\]. We can see this better if we extract the values manually and make our own plot.

```{r}
x1 <- pd1[["param.x"]]
y1 <- pd1[["posterior"]]
df <- data.frame(x1, y1)
df |> 
  ggplot(aes(x = x1, y = y1))+
  geom_area(fill="orange", alpha=.5)+
  xlim(c(0,7))
```

We can perhaps trucate the posterior, like the figure above, or we could use a discrete prior

`mu` is actually not discrete, just the data are. But with this approach, mu is also discrete, but we can add many steps to make it core or less continous. I use 100 steps below.

```{r}
mu <- seq(1, 7, length.out=100)
```

```{r}
pd2 <- normdp(x = random1,
       sigma.x = national.sd,
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))

```

OK, but there is practically zero change of mu being 7. We only have one sample. Shouldn't there be a wider spread of probabilities?

Actually, the probability that mu is 7 is calculated to be

```{r}
min(pd2$posterior)
```

So not zero, but still very small. And if the st.dev is 2, then perhaps this is correct.

```{r}
dnorm(7, mean=random1, sd=national.sd)
```

This is also a small number, although not identical. I'm not sure I understand the maths here.

Let's try with *n* = 10.

```{r}
pd10 <- normdp(x = random10,
       sigma.x = national.sd,
       n.mu = 50, # not having an effect
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))
```

Lets try some more realistic samples.

```{r}
sampleHigh2 <- round(runif(2, 5,7))
sampleHigh5 <- round(runif(5, 5,7))
sampleHigh10 <- round(runif(10, 5,7))
sampleHigh20 <- round(runif(20, 5,7))
```

```{r, output=F}
pd_h2 <- normdp(x = sampleHigh2,
       sigma.x = national.sd,
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))

pd_h5 <- normdp(x = sampleHigh5,
       sigma.x = national.sd,
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))

pd_h10 <- normdp(x = sampleHigh10,
       sigma.x = national.sd,
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))

pd_h20 <- normdp(x = sampleHigh20,
       sigma.x = national.sd,
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))
```

```{r}
data.frame(x = mu,
                 "n2" = pd_h2$posterior,
                 "n5" = pd_h5$posterior,
                 "n10" = pd_h10$posterior,
                 "n20" = pd_h20$posterior) |>
  pivot_longer(!x, 
               names_to = "n") |>
  mutate(n2 = str_remove(n, "n"),
         n3 = as.numeric(n2),
         n = fct_reorder(n2, n3)) |>
  ggplot(aes(x = x, 
             y = n, 
             height = value,
             fill = n))+
  geom_density_ridges(alpha=.5,
                      stat = "identity")+
  theme_ridges()

```

This is looking good so far.

## Weighted mean

Now for the next bit. We actually need the weighted mean for the population.

Let's use `pd10` to compare against. `pd10` used the `random10` as the population sample.

```{r}
plot(pd10)
```

We want to add weights to `random10`. Creating vectors of weights that down weigh the low numbers:

```{r}
random10_w <- random10
random10_w2 <- random10^2
```

I have custumised `Bolstad::normdp` to handle weighted means with a function called `wgt_mean`.

```{r}
source("../bayes_weightedMean.R")
```

```{r, output = F}
dw1 <- wgt_mean(x = random10,
         weights = random10_w, #notice new argument
       sigma.x = national.sd,
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))

dw2 <- wgt_mean(x = random10,
         weights = random10_w2,
       sigma.x = national.sd,
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))
```

```{r}
data.frame(x = mu,
          "No_weights"    = pd10$posterior,
          "Proportional" = dw1$posterior,
          "Exponential"  = dw2$posterior) |>
  pivot_longer(!x, 
               names_to = "Weights") |>
  mutate(Weights = fct_relevel(Weights, 
                               "No_weights",
                               "Proportional",
                               "Exponential")) |>
    ggplot(aes(x = x, 
             y = Weights, 
             height = value,
             fill = Weights))+
  geom_density_ridges(alpha=.5,
                      stat = "identity")+
  theme_ridges()
```

Our custom function has succesfully increased the weight on the high values in the sample.

How about if the sample size varies.

```{r, output=F}
dw3 <- wgt_mean(x = sampleHigh2,
         weights = sampleHigh2,
       sigma.x = national.sd,
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))

dw4 <- wgt_mean(x = sampleHigh10,
         weights = sampleHigh10,
       sigma.x = national.sd,
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))

dw5 <- wgt_mean(x = sampleHigh20,
         weights = sampleHigh20,
       sigma.x = national.sd,
       mu = mu,
       mu.prior = rep(1/length(mu), times=length(mu)))
```

```{r}
data.frame(x = mu,
                 "n2" = dw3$posterior,
                 "n10" = dw4$posterior,
                 "n20" = dw5$posterior) |>
  pivot_longer(!x, 
               names_to = "n") |>
  mutate(n2 = str_remove(n, "n"),
         n3 = as.numeric(n2),
         n = fct_reorder(n2, n3)) |>
  ggplot(aes(x = x, 
             y = n, 
             height = value,
             fill = n))+
  geom_density_ridges(alpha=.5,
                      stat = "identity")+
  theme_ridges()
```

It works as predicted.