### 2024-02-28 manuscript/Appendix A.qmd
Appendix A

---
title: Appendix A
subtitle: For the article 'Using spatially biased variables in ecosystem condition accounting with a GIS based workflow'
author:
  - name: Anders Lorentzen Kolstad
    email: anders.kolstad@nina.no
    orchid: https://orcid.org/0000-0002-9623-9491
    affiliations: 
        - id: nina
          name: Norwegian Institute for Nature Research
          department: Department of Terrestrial Ecology
          address: Pb 5685 Torgarden
          city: Trondheim
          postal-code: 7485
  - name: Matthew Grainger
    email: matthew.grainger@nina.no
    orchid: https://orcid.org/0000-0001-8426-6495
    affiliations:
        - ref: nina
date: last-modified
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Ecological Indicators
      formatting: preprint
      model: 3p
      cite-style: authoryear
    include-in-header:
      text: '\usepackage{lineno}\linenumbers'
editor: 
  markdown: 
    wrap: sentence
crossref:
  fig-title: |
    Supplementary Figure
  fig-prefix: Fig S.
  tbl-title: |
    Supplementary Table
  tbl-prefix: Fig S.
execute:
  echo: false
---

```{r fig-studyLocations}
#| fig.cap: "Position of the three example municiaplieies in Norway."
#| out.width: '50%'

knitr::include_graphics("../images/positionMap.jpg")
```

{{< pagebreak >}}
```{r tbl-EEA}
#| tbl-cap: "Idicator values for three ecosystem condition indicators in three 
#| Norwegian municipalities. Values are medians with 2.5 and 97.5 percentiles
#| for the distribution of possible values."

(tbl <- readRDS("../output/EEA-tbl.RDS"))
```