### 2024-02-29 manuscript/Appendix A.qmd
workflow example

---
title: Appendix A
subtitle: For the article 'Using spatially biased variables in ecosystem condition accounting with a GIS based workflow'
author:
  - name: Anders Lorentzen Kolstad
    email: anders.kolstad@nina.no
    orchid: https://orcid.org/0000-0002-9623-9491
    affiliations: 
        - id: nina
          name: Norwegian Institute for Nature Research
          department: Department of Terrestrial Ecology
          address: Pb 5685 Torgarden
          city: Trondheim
          postal-code: 7485
  - name: Matthew Grainger
    email: matthew.grainger@nina.no
    orchid: https://orcid.org/0000-0001-8426-6495
    affiliations:
        - ref: nina
date: last-modified
format:
  pdf
editor: 
  markdown: 
    wrap: sentence
crossref:
  fig-title: |
    Figure S
  tbl-prefix: Fig S.
  fig-prefix: Fig S.
  tbl-title: |
    Table S
execute:
  echo: false
---

```{r fig-studyLocations}
#| fig.cap: "Position of the three example municiaplieies in Norway."
#| out.width: '50%'

knitr::include_graphics("../images/positionMap.jpg")
```

```{r fig-studyLocations}
#| fig.cap: "Barplot showing the total area from each homogeneous impact area 
#| in three Norwegian municipalities"
#| out.width: '50%'

knitr::include_graphics("../images/infra-dist-plot.jpg")
```

```{r fig-scaling}
#| eval: true
#| include: true
#| out.width: '70%'
#| fig.cap:
#| - "Custom re-scaling of variables to indicators. Unique threshold values are 
#| assignes to each variable and this value is normalised to become 0.6 on the 
#| indicator scale (y-axis). The variables also differ in the number of possible 
#| values they can take, hence the varying number of points for the three panes."

knitr::include_graphics("../images/scaling-plot.jpg")
```

{{< pagebreak >}}
```{r tbl-EEA}
#| tbl-cap: "Indicator values for three ecosystem condition indicators in three 
#| Norwegian municipalities. Values are medians with 2.5 and 97.5 percentiles
#| for the distribution of possible values."

(tbl <- readRDS("../output/EEA-tbl.RDS"))
```