### 2024-03-18 R/presentation.qmd
add presentation

---
title: "Presenation"
format: revealjs
editor: visual
include-in-header: 
  text: |
    <style>
      .custom-small table {
        font-size: .5em
      }
    </style>
---

## Horizontal aggregation of spatially biased ecosystem condition indicators – a GIS based workflow

Anders L. Kolstad

Mathew Grainger

Marianne Evju

<br> *1st draft* - aimed at Ecosystem Indicators

## Question

-   Is there a way to make use of spatially biased data in ecosystem condition indicators?\

::: {.callout-note appearance="minimal"}
Ecosystem condition indicators should be wall-to-wall. For municipalities with little available data, the indicators values could be inferred from projections based on a model. If municipalities are somehow judged based on their indicator scores, it is not ideal if this is based on an assumption that all municipality are the same.
:::

-   Can we make local wall-to-wall indicator maps from sporadic data without relying on data from other municipalities?

## Workflow

```{r}
#| fig.cap: ""
knitr::include_graphics("images/aggregation-workflow.jpg")
```

## Variables

::: custom-small
| id  | Variable code        | Variable name                                    |
|---------------|---------------|-------------------------------------------|
| 1   | 7FK                  | Prevalence of alien species                      |
| 2   | 7SE                  | Human caused abration or abration-caused erosion |
| 3   | PRSL                 | *as above*                                       |
| 4   | 7TK                  | Tracks from large vehicles                       |
| 5   | PRTK                 | *as above*                                       |
| 6   | 7GR-GI               | Trenching intensity                              |
| 7   | Infrastructure Index | Infrastructure Index                             |
| 8   | HIA                  | Homogeneous Impact Area                          |
:::

## Workflow 2

![](images/workflow-01.jpg)

## Pilot municipalities

![](images/positionMap.jpg)

## Pilot municipalities

![](images/studyLocations.jpg)

## Scaling![](images/scaling-plot.jpg)

## Trenching variable

![](images/indicator-magnify.jpg)

## Validation

![](images/validation-plot.jpg)

## Generalise

![](images/forest-plot.jpg)

## Spread

![](images/spread-na-small.jpg)

## Aggregate

![](images/ridgePlot.jpg)